---
title: "STAT2170_Assignment 1_45941866"
author: "Olivia Paleologos"
date: "16/05/2021"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
surg <- read.csv("~/Desktop/Assignment datasets-20210516/surg.dat", sep="")
```
# Question 1

A medical research team wants to investigate the survival time of patients that have a particular type of liver operation as part of their treatment. For each patient in the study, the following variables were recorded.

# a. Produce a scatterplot of the data and comment on the features of the data and possible relationships between the response and predictors and relationships between the predictors themselves.

The scatterplot that is shown below, is describing a negative linear regression. Showcasing 3 outliers that is above 5. The cluster is shown declining on the x value and increasing on the y value which determines it a negative scatterplot. This is also showing the relationship of the predictors which is highlighting that maybe H1 is accepted and H0 is rejected.

```{r chunk1}
x <- mtcars$wt
y <- mtcars$mpg
plot(x, y, main = "surge")
```

# b.Compute the correlation matrix of the dataset and comment.

```{r chunk2}
head(surg)
surg.1 = lm(liver ~ prognosis + enzyme + age + survival, data = surg)
summary(surg.1)
```

Through the table above, it can be seen that b5 is the most significant, this is shown by looking at the p-value 0.000748 *** is is <0.05.

# c.Fit a model using all the predictors to explain the survival response. 


• Write down the Hypotheses for the Overall ANOVA test of multiple regression.

H0: no relationship between the response and predictors and relationships between the predictors themselves.

H1: there is a relationship between the response and predictors and relationships between the predictors themselves.

```{r chunk3}
mod = lm(survival ~ age + enzyme + prognosis, data = surg)
anova(mod)
```

• Produce an ANOVA table for the overall multiple regression model (One combined regression SS source is sufficient).

Now, this anova table is showing the significant variables of enzyme and prognosis. 
Due to the change of the p-values it is showing how these variables considered have evidence of importance with the relation of survival.

```{r chunk3.1}
qf(.95, df1 = 1, df2 = 50)
```
• Compute the F statistic for this test. 

As F-test is one-tailed we only care about a larger variance and in this case the F-stat is 4.03431 which is considered large. 


• Compute the P-Value

The P-value which is shown as 1.048e-06 in the multiple regression table.

• State your conclusion (both statistical conclusion and contextual conclusion).

My conclusion is that due to the p-value being significant, it is showcasing that we reject H0 and accept H1.


# d.Using model selection procedures discussed in the course, find the best multiple regression model that explains the data.

```{r chunk4}
surg.3 = lm(survival ~ prognosis + enzyme, data = surg)
summary(surg.3)
```
The multiple regression table above is indicating that survival and its relation to prognosis and enzyme has shown that it is significant though looking at their p-values.

I have cut age and liver out of the data due to their non-significant values.
And as doing so, it has picked up that prognosis and enzyme is very significant.


# e.Validate your final model and comment why it is not appropriate to use the multiple regression model to explain the survival time.

```{r chunk5}
par(mfrow = c(1,2))
plot(surg.3, which = 1:2)
plot(resid(surg.3) ~ survival + prognosis + enzyme, data = surg)
```
The Normal Q-Q plot is highlighting a negative linear regression and indicating 3 outliers. The Residuals vs Fitted plot is indicating a bell curve in the linear regression with one outlier.

# f.Re-fit the model using log(survival) as the new response variable. In your answer,

Use the model selection procedure discussed in the course starting with log(survival) as the response and start with all the predictors.

```{r chunk6}
surg.3 = lm(survival ~ prognosis + enzyme, data = surg)
summary(surg.3)$coef
```

# g.Validate your final model with the log(survival) response. In particular, in your answer,

Explain why the regression model with log(survival) response variable is superior to the model with the survival response variable

# Question 2

A car manufacturer wants to study the fuel efficiency of a new car engine. It wishes to account for any differences between the driver and production variation. The manufacturer randomly selects 5 cars from the production line and recruits 4 different test drivers.

# a. For this study, is the design balanced or unbalanced? Explain why.

In ANOVA and Design of Experiments, a balanced design has an equal number of observations for all possible level combinations. This is compared to an unbalanced design, which has an unequal number of observations. 

```{r chunk7}
kml <- read.csv("~/Desktop/Assignment datasets-20210516/kml.dat", sep="")
```

# b. Construct two different preliminary graphs that investigate different features of the data and comment.

```{r chunk8}
summary(kml)
par(mfrow = c(1, 2))
plot(kml)
```

This first plot is showing the relationship between kmL to driver and car, driver to car and kmL and car to driver and kmL. 

```{r plot2}
x <- mtcars$wt
y <- mtcars$mpg
plot(x, y, main = "kml")
```

This scatterplot is highlighting a positive linear regression with three outliers. These outliers would suggest that it is an unbalanced data experiment. 


# c. Analyse the data, stating null and alternative hypothesis for each test, and check assumptions.

Null hypothesis: No difference between the driver and the production variation
Hypothesis 2: There is a difference between the driver and the production variation

To check assumptions a regression model will be implemented.

```{r chunk9}
mod.1 = lm(kmL~ driver + car, data = kml)
anova(mod.1)
```

# d. State your conclusions about the effect of driver and car on the efficiency kmL. 

Through running an anova multiple regression model it is showing how from the efficiency kmL to the effect of driver and car the p-values of driver and car are showing to be significant as they are both <0.05. This is indicating that we reject the null hypothesis and accept the new hypothesis.

